---
title: "ABC Transporters"
author: "Daniel Cook"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cosmo
    toc: yes
---

#### Blast Command

```{r, cache = F, echo = F}
library(dplyr)
library(readr)
library(tidyr)
library(stringr)
library(reutils)
library(XML)
library(cowplot)
library(DT)
library(GenomicRanges)

setwd("/Users/dancook/Dropbox/Andersenlab/LabFolders/Dan/Projects/abc/")

fasta <- "Sap.fasta"
db <- "reference/all_bacteria"

query_name <- paste0("blast_results/", fasta, ".blast.txt")
comm <- paste("export BLASTDB=/usr/local/share/blast; tblastn -query ", fasta,
              "-db", db, 
              "-task tblastn",
              "-max_target_seqs 20000",
              "-outfmt '6  sscinames scomnames staxids qseqid sseqid pident length mismatch gapopen qstart qend sstart send evalue bitscore'",
              ">",
              query_name)
print(comm)
```

```{r blast_search, echo = F, cache = T}
#system(comm)


results <- read_tsv( query_name, col_names = c("Species",
                                       "Name",
                                       "TaxID",
                                       "QueryID",
                                       "SubjectID",
                                       "Percent_Identity",
                                       "Alignment_Length",
                                       "Mismatches",
                                       "Gap_Openings",
                                       "Q.Start",
                                       "Q.End",
                                       "S.Start",
                                       "S.End",
                                       "E",
                                       "Bits") )  %>%
separate(SubjectID, into = c("name_drop", "gi", "ref_drop","accession"), sep = "\\|", extra = "drop", convert = T) %>%
dplyr::select(-name_drop, -ref_drop) %>%
dplyr::rename(POS_Start = S.Start, POS_End = S.End) %>%
dplyr::mutate(CHROM = str_match(Name, "chromosome ([A-Za-z0-9])")[,2]) %>%
dplyr::select(Species, Name, TaxID, QueryID, CHROM, POS_Start, POS_End, accession,  everything()) 


save(results, file = "bacteria_blast_results.Rdata")

load("blast_hits.RData")
trim <- function (x) gsub("^\\s+|\\s+$", "", x)


genomes <- read_delim("reference/genomes.txt", "|", col_names = c("gi", "ref", "accession", "name")) %>%
           select(-ref) %>%
           mutate_each(funs(trim), everything()) %>%
           mutate(gi = as.integer(gi))


df <- r %>% ungroup() %>% 
             left_join(genomes, by = c("gi", "accession")) %>%
             filter(!is.na(QueryID)) %>%
             mutate(Species = trim(Species)) %>%
             group_by(accession) %>%
             filter(any("SapA" %in% QueryID)) %>%
             mutate(direction = ifelse(S.End > S.Start, "+", "-")) %>%
             mutate(swap1 = ifelse(S.Start > S.End, S.End, S.Start)) %>%
             mutate(swap2 = ifelse(S.Start > S.End, S.Start, S.End)) %>%
             mutate(S.Start = swap1, S.End = swap2) %>%
             select(-Species, -Name, -TaxID, -gi, -swap1, -swap2)) %>%
             select(accession, S.Start, S.End, direction,  everything())


df2 <- filter(df, QueryID == "SapA") %>%
           mutate(S.Start = S.Start-7000, S.End = S.End+7000, SapA_Start = S.Start) %>%
           select(accession, S.Start, S.End, SapA_Start) %>%
           dplyr::rename(SapA.Start_region=S.Start, SapA.End_region=S.End) %>%
           left_join(df, by = c("accession")) %>%
           filter( (S.Start > SapA.Start_region & S.Start < SapA.End_region) ) %>%
           group_by(accession, SapA_Start, QueryID, S.Start, S.End) %>%
           distinct() %>%
           ungroup() %>%
           group_by(accession, SapA_Start) %>%
           filter(n() >= 3) %>%
           separate(name, sep = ",", into = c("Species", "Contig"), remove = F, extra = "drop") %>%
           mutate(Species = trim(str_replace(Species,"chromosome",""))) %>%
           mutate(Species = trim(str_replace(Species,"complete genome",""))) %>%
           group_by(accession, SapA_Start) %>%         
           mutate(SapA = sum(QueryID == "SapA"),
                  SapB = sum(QueryID == "SapB"),
                  SapC = sum(QueryID == "SapC"),
                  SapD = sum(QueryID == "SapD"),
                  SapF = sum(QueryID == "SapF"),
                  SapZ = sum(QueryID == "SapZ"),
                  mean_percent = mean(Percent_Identity)) %>%
          dplyr::ungroup() %>%
                  SapZ = sum(QueryID == "SapZ"), 
                  SapD_only = sum(QueryID == "SapD_only"),
                  SapF_only = sum(QueryID == "SapF_only"),
          dplyr::arrange(desc(mean_percent)) %>%
                  SapD_nbd  = sum(QueryID == "SapD_nbd"),
                  SapF_nbd  = sum(QueryID == "SapF_nbd")) %>%
          dplyr::filter(SapA > 0, (SapB > 0 | SapC > 0), (SapD > 0 | SapF > 0) ) %>%
          dplyr::mutate(gene_string = paste0("gs", SapA, SapB, SapC, SapD, SapF, SapZ))



df2$operon_id <- group_indices(df2)


```

### Distribution of ABC operons by species

```{r Distribution of N Operons, echo = F}

distr_abc <- select(df2, operon_id, accession, Species) %>%
  group_by(Species) %>%
  summarize(n_abc = n())

ggplot(distr_abc, aes(x = n_abc)) +
  geom_histogram(binwidth = 1, fill = "#FF7F50") +
  scale_y_continuous(expand = c(0,0), limits = c(0, 200)) +
  scale_x_continuous(expand = c(0,0)) +
  labs(x = "Number of ABC Operons Identified in a species", y = "Frequency") +
  theme_bw()

```

__Figure__ - Number of operons identified in each species. The only filter implemented removes operons with less than 3 genes.

* __Total number of species__ - `r select(distr_abc, Species) %>% nrow()`
* __Max__ - `r max(distr_abc$n_abc)`
* __Median__ - `r median(distr_abc$n_abc)`
* __Mean__ - `r mean(distr_abc$n_abc)`
* __Std Dev__ - `r sd(distr_abc$n_abc)`
* __Min__ - `r min(distr_abc$n_abc)`


`r datatable( distr_abc %>% arrange(desc(n_abc)), style = "bootstrap", options = list(pageLength = 10), class = 'cell-border stripe')`

### Distribution of % Identity

```{r Percent_Identities}


ggplot(df2, aes(x = Percent_Identity, fill = QueryID)) +
  geom_histogram(binwidth = 0.5) +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0), limits = c(0,100)) +
  labs(x = "% Identity", y = "Frequency") +
  theme_bw() +
  facet_grid(QueryID~.)


```

### Filter Further

```{r}

df3 <- df2 %>% dplyr::filter(!(QueryID %in% c("SapD_only", "SapF_only", "SapD_nbd", "SapF_nbd"))) %>%
        dplyr::arrange(QueryID) %>%
        dplyr::mutate(gap = dplyr::lead(S.Start) - S.End) %>%
        dplyr::select(operon_id, accession, S.Start, S.End, direction, gap, everything()) %>%
        dplyr::filter(gap > 0 | is.na(gap)) %>%
        dplyr::mutate(gap =  lead(S.Start) - S.End) %>%  
        dplyr::filter(gap > 0 | is.na(gap)) %>%
        dplyr::mutate(gap =  lead(S.Start) - S.End) %>% 
          dplyr::filter(gap > 0 | is.na(gap)) %>%
        dplyr::mutate(gap =  lead(S.Start) - S.End) %>% 
          dplyr::filter(gap > 0 | is.na(gap)) %>%
        dplyr::mutate(gap =  lead(S.Start) - S.End) %>% 
        dplyr::group_by(operon_id) %>%
           mutate(SapA = sum(QueryID == "SapA"),
                  SapB = sum(QueryID == "SapB"),
                  SapC = sum(QueryID == "SapC"),
                  SapD = sum(QueryID == "SapD"),
                  SapF = sum(QueryID == "SapF"),
                  SapZ = sum(QueryID == "SapZ")) %>%
        dplyr::filter(n() >= 3)
 
freq <- df3 %>% dplyr::ungroup() %>% group_by(SapA, SapB, SapC, SapD, SapF, SapZ) %>%
  summarize(n = n())
lapply(1:max(df2$operon_id), function(index) {
  
  x <- unique(df2$operon_id)[[index]]
  
  plot_df <- df2 %>%
    dplyr::filter(operon_id == x) %>%
    dplyr::mutate(row = row_number()) %>%
    dplyr::mutate(title =  paste0(species, "(gi=", gi,"); E= ", E, sep = " "))
  
  ggplot(plot_df %>% dplyr::filter(operon_id == x)) +
    geom_rect(aes(xmin = S.Start, xmax = S.End, ymin = row-1, ymax = row, color=QueryID, fill = QueryID, alpha = Percent_Identity)) +
    geom_text(aes(x = (S.Start + ((S.End - S.Start)/2)), y = row-0.5, label = paste0(QueryID, " (", Percent_Identity, ") "))) +
    labs(title = plot_df$title[[1]]) +
    facet_grid(operon_id ~ ., space = "free_x", scales = "free") +
    theme_bw() +
    theme(axis.text.y  = element_blank(),
          axis.title.x  = element_blank(),
          axis.title.y = element_blank(),
          axis.ticks.y = element_blank(),
          legend.position = "None")
  
  ggsave(paste0("plots/", sprintf("%05d", index), "_", plot_df$gene_string[[1]], "_", plot_df$accession[[1]],".svg"), height = 3, width = 10, limitsize = F )
})

```
